Personality and political beliefs across the lifespan

### 2019-12-11 scripts/manuscript.Rmd
added intro

```{r table1, echo = FALSE, results = "asis"}
#this table is of the mean personality trait scores by social conservatism scores
tidy_table_sc <- tidy %>% 
  filter(conservatism_type == 'Social conservatism') %>%
  mutate(conservatism_score = case_when(       
      conservatism_score <= 19 & conservatism_score >= 0 ~ "[0-19]",
      conservatism_score <= 39 & conservatism_score >= 20 ~ "[20-39]",
      conservatism_score <= 59 & conservatism_score >= 40 ~ "[40-59]",
      conservatism_score <= 79 & conservatism_score >= 60 ~ "[60-79]",
      TRUE ~ "[80-100]"))
tidy_table_sc <- tidy_table_sc %>%
  group_by(personality_trait, conservatism_score) %>% 
  summarize(mean_personality_score = mean(personality_score)) %>% 
  pivot_wider(names_from = conservatism_score, values_from = mean_personality_score)

```{r linearmodels_conservatism_ocean_tables, echo = FALSE, results = "asis"}
apa_sc_ocean <- apa_print(sc_ocean)
apa_sc_ocean$table %>% 
  apa_table(caption = "Regression Table Predicting Social Conservatism From Big-Five Personality Traits.", note = "Residual standard error: 25.13 on 125 degrees of freedom.
Multiple R-squared: 0.168, Adjusted R-squared: 0.135. F(5, 125): 5.051, p-value: 0.0003.", escape = FALSE)

apa_ec_ocean <- apa_print(ec_ocean)
apa_ec_ocean$table %>% 
  apa_table(caption = "Regression Table Predicting Economic Conservatism From Big-Five Personality Traits.", note = "Residual standard error: 21.3 on 125 degrees of freedom. Multiple R-squared: 0.081, Adjusted R-squared: 0.044. F(5, 125): 2.207, p-value: 0.058.", escape = FALSE)
```

# Methods

```{r, include = FALSE}
#Age statistics
mean_age <- mean(as.numeric(data$age)) %>% round(2)
sd_age <- sd(as.numeric(data$age)) %>% round(2)
r1_age <- range(as.numeric(data$age))[1]
r2_age <- range(as.numeric(data$age))[2]

@@

<div id = "refs"></div>

>>>>

The Five Factor Model is a widely used framework for measuring and classifying personality into five dimensions [@goldberg1992development; @mccrae1999age; @john1999big]. Those who score high in *openness to experiences* tend to describe themselves as curious, deep thinkers compared to those who score low in openness to experience, who describe themselves as traditional and conventional. Those who score high in *conscientiousness* are generally reliable and hardworking, while those who score low in conscientiousness tend to be disorganized and impulsive. Those high in *extraversion* are talkative and energetic, while those high in introversion are reserved and quiet. Those who score high in *agreeableness* describe themselves as kind and trusting, while those who score low in agreeableness say they are quarrelsome and critical. And lastly, those who are high in *emotional stability* are calm individuals who handle stress well, while those high in neuroticism tend to be anxious individuals who are easily upset.

Given that these personality traits have been found to predict a multitude of cognitions and behaviors, from well-being to cognitive abilities [@hayes2003big; @curtis2015relationship; @barlett2012direct], previous studies have looked at whether personality can predict political beliefs and affiliation [@jost2009personality; @hirsh2010compassionate]. The most robust findings are that openness to experience is positively correlated with liberalness [@carney2008secret; @ekehammar2004matters], while conscientiousness is positively correlated with conservativeness [@stenner2005rights; @barbaranelli2007voters].

This project uses data from a larger study that measured political beliefs and the big five personality traits. Consistent with prior research, we hypothesize that we will find a positive relation between conscientiousness and conservatism and a negative relation between openness to experience and conservatism. We will first make exploratory tables and plots that explore the relation between all big five personality traits and social and economic conservatism, and then we will run two linear models with traits as predictors and types of conservatism as outcome variables.

# Method

```{r, include = FALSE}
#I'm am going to use these descriptives in the participants section
#Age statistics
mean_age <- mean(as.numeric(data$age)) %>% round(2)
sd_age <- sd(as.numeric(data$age)) %>% round(2)
r1_age <- range(as.numeric(data$age))[1]
r2_age <- range(as.numeric(data$age))[2]

@@

<div id = "refs"></div>
```{r table1, echo = FALSE, results = "asis"}
#this table is of the mean personality trait scores by social conservatism scores
tidy_table_sc <- tidy %>% 
  filter(conservatism_type == 'Social conservatism') %>%
  mutate(conservatism_score = case_when(       
      conservatism_score <= 19 & conservatism_score >= 0 ~ "[0-19]",
      conservatism_score <= 39 & conservatism_score >= 20 ~ "[20-39]",
      conservatism_score <= 59 & conservatism_score >= 40 ~ "[40-59]",
      conservatism_score <= 79 & conservatism_score >= 60 ~ "[60-79]",
      TRUE ~ "[80-100]"))
tidy_table_sc <- tidy_table_sc %>%
  group_by(personality_trait, conservatism_score) %>% 
  summarize(mean_personality_score = mean(personality_score)) %>% 
  pivot_wider(names_from = conservatism_score, values_from = mean_personality_score)

```{r table3, echo = FALSE, results = "asis"}
apa_sc_ocean <- apa_print(sc_ocean)
apa_sc_ocean$table %>% 
  apa_table(caption = "Regression Table Predicting Social Conservatism From Big-Five Personality Traits.", note = "Residual standard error: 25.13 on 125 degrees of freedom.
Multiple R-squared: 0.168, Adjusted R-squared: 0.135. F(5, 125): 5.051, p-value: 0.0003.", escape = FALSE)
```

```{r table4, echo = FALSE, results = "asis"}
apa_ec_ocean <- apa_print(ec_ocean)
apa_ec_ocean$table %>% 
  apa_table(caption = "Regression Table Predicting Economic Conservatism From Big-Five Personality Traits.", note = "Residual standard error: 21.3 on 125 degrees of freedom. Multiple R-squared: 0.081, Adjusted R-squared: 0.044. F(5, 125): 2.207, p-value: 0.058.", escape = FALSE)
```

### 2019-12-11 scripts/manuscript.tex
added intro

\begin{table}[tbp]
\begin{center}
\begin{threeparttable}
\caption{\label{tab:linearmodels_conservatism_ocean_tables}Regression Table Predicting Social Conservatism From Big-Five Personality Traits.}
\begin{tabular}{lllll}
\toprule
Predictor & \multicolumn{1}{c}{$b$} & \multicolumn{1}{c}{95\% CI} & \multicolumn{1}{c}{$t(125)$} & \multicolumn{1}{c}{$p$}\\
\midrule
Intercept & 25.82 & $[0.07$, $51.58]$ & 1.98 & .049\\
Openness to experiences & -3.73 & $[-7.50$, $0.04]$ & -1.96 & .052\\
Conscientiousness & 0.07 & $[-3.98$, $4.11]$ & 0.03 & .974\\
Extraversion & 1.56 & $[-1.42$, $4.55]$ & 1.04 & .302\\
Agreeableness & 3.93 & $[0.69$, $7.17]$ & 2.40 & .018\\
Emotional stability & 4.30 & $[0.70$, $7.91]$ & 2.37 & .020\\
\bottomrule
\addlinespace
\end{tabular}
\begin{tablenotes}[para]
\normalsize{\textit{Note.} Residual standard error: 25.13 on 125 degrees of freedom.
Multiple R-squared: 0.168, Adjusted R-squared: 0.135. F(5, 125): 5.051, p-value: 0.0003.}
\end{tablenotes}
\end{threeparttable}
\end{center}
\end{table}

\begin{table}[tbp]
\begin{center}
\begin{threeparttable}
\caption{\label{tab:linearmodels_conservatism_ocean_tables}Regression Table Predicting Economic Conservatism From Big-Five Personality Traits.}
\begin{tabular}{lllll}
\toprule
Predictor & \multicolumn{1}{c}{$b$} & \multicolumn{1}{c}{95\% CI} & \multicolumn{1}{c}{$t(125)$} & \multicolumn{1}{c}{$p$}\\
\midrule
Intercept & 56.54 & $[34.71$, $78.37]$ & 5.13 & < .001\\
Openness to experiences & -3.79 & $[-6.98$, $-0.59]$ & -2.35 & .021\\
Conscientiousness & 1.78 & $[-1.65$, $5.20]$ & 1.03 & .307\\
Extraversion & 0.61 & $[-1.92$, $3.14]$ & 0.48 & .632\\
Agreeableness & -1.13 & $[-3.88$, $1.61]$ & -0.82 & .415\\
Emotional stability & 2.34 & $[-0.71$, $5.39]$ & 1.52 & .131\\
\bottomrule
\addlinespace
\end{tabular}
\begin{tablenotes}[para]
\normalsize{\textit{Note.} Residual standard error: 21.3 on 125 degrees of freedom. Multiple R-squared: 0.081, Adjusted R-squared: 0.044. F(5, 125): 2.207, p-value: 0.058.}
\end{tablenotes}
\end{threeparttable}
\end{center}
\end{table}

\section{Methods}\label{methods}

@@

@@

@@

@@

@@

@@

@@

@@

>>>>

The Five Factor Model is a widely used framework for measuring and
classifying personality into five dimensions (Goldberg, 1992; John \&
Srivastava, 1999; McCrae et al., 1999). Those who score high in
\emph{openness to experiences} tend to describe themselves as curious,
deep thinkers compared to those who score low in openness to experience,
who describe themselves as traditional and conventional. Those who score
high in \emph{conscientiousness} are generally reliable and hardworking,
while those who score low in conscientiousness tend to be disorganized
and impulsive. Those high in \emph{extraversion} are talkative and
energetic, while those high in introversion are reserved and quiet.
Those who score high in \emph{agreeableness} describe themselves as kind
and trusting, while those who score low in agreeableness say they are
quarrelsome and critical. And lastly, those who are high in
\emph{emotional stability} are calm individuals who handle stress well,
while those high in neuroticism tend to be anxious individuals who are
easily upset.

Given that these personality traits have been found to predict a
multitude of cognitions and behaviors, from well-being to cognitive
abilities (Barlett \& Anderson, 2012; Curtis, Windsor, \& Soubelet,
2015; Hayes \& Joseph, 2003), previous studies have looked at whether
personality can predict political beliefs and affiliation (Hirsh,
DeYoung, Xu, \& Peterson, 2010; Jost, West, \& Gosling, 2009). The most
robust findings are that openness to experience is positively correlated
with liberalness (Carney, Jost, Gosling, \& Potter, 2008; Ekehammar,
Akrami, Gylje, \& Zakrisson, 2004), while conscientiousness is
positively correlated with conservativeness (Barbaranelli, Caprara,
Vecchione, \& Fraley, 2007; Stenner, 2005).

This project uses data from a larger study that measured political
beliefs and the big five personality traits. Consistent with prior
research, we hypothesize that we will find a positive relation between
conscientiousness and conservatism and a negative relation between
openness to experience and conservatism. We will first make exploratory
tables and plots that explore the relation between all big five
personality traits and social and economic conservatism, and then we
will run two linear models with traits as predictors and types of
conservatism as outcome variables.

\section{Method}\label{method}

@@

\hypertarget{ref-barbaranelli2007voters}{}
Barbaranelli, C., Caprara, G. V., Vecchione, M., \& Fraley, C. R.
(2007). Voters' personality traits in presidential elections.
\emph{Personality and Individual Differences}, \emph{42}(7), 1199--1208.

\hypertarget{ref-barlett2012direct}{}
Barlett, C. P., \& Anderson, C. A. (2012). Direct and indirect relations
between the big 5 personality traits and aggressive and violent
behavior. \emph{Personality and Individual Differences}, \emph{52}(8),
870--875.

@@

\hypertarget{ref-carney2008secret}{}
Carney, D. R., Jost, J. T., Gosling, S. D., \& Potter, J. (2008). The
secret lives of liberals and conservatives: Personality profiles,
interaction styles, and the things they leave behind. \emph{Political
Psychology}, \emph{29}(6), 807--840.

@@

\hypertarget{ref-curtis2015relationship}{}
Curtis, R. G., Windsor, T. D., \& Soubelet, A. (2015). The relationship
between big-5 personality traits and cognitive ability in older
adults--a review. \emph{Aging, Neuropsychology, and Cognition},
\emph{22}(1), 42--71.

\hypertarget{ref-ekehammar2004matters}{}
Ekehammar, B., Akrami, N., Gylje, M., \& Zakrisson, I. (2004). What
matters most to prejudice: Big five personality, social dominance
orientation, or right-wing authoritarianism? \emph{European Journal of
Personality}, \emph{18}(6), 463--482.

@@

\hypertarget{ref-goldberg1992development}{}
Goldberg, L. R. (1992). The development of markers for the big-five
factor structure. \emph{Psychological Assessment}, \emph{4}(1), 26.

@@

\hypertarget{ref-hayes2003big}{}
Hayes, N., \& Joseph, S. (2003). Big 5 correlates of three measures of
subjective well-being. \emph{Personality and Individual Differences},
\emph{34}(4), 723--727.

@@

\hypertarget{ref-hirsh2010compassionate}{}
Hirsh, J. B., DeYoung, C. G., Xu, X., \& Peterson, J. B. (2010).
Compassionate liberals and polite conservatives: Associations of
agreeableness with political ideology and moral values.
\emph{Personality and Social Psychology Bulletin}, \emph{36}(5),
655--664.

\hypertarget{ref-john1999big}{}
John, O. P., \& Srivastava, S. (1999). The big five trait taxonomy:
History, measurement, and theoretical perspectives. \emph{Handbook of
Personality: Theory and Research}, \emph{2}(1999), 102--138.

\hypertarget{ref-jost2009personality}{}
Jost, J. T., West, T. V., \& Gosling, S. D. (2009). Personality and
ideology as determinants of candidate preferences and ``obama
conversion'' in the 2008 us presidential election. \emph{Du Bois Review:
Social Science Research on Race}, \emph{6}(1), 103--124.

\hypertarget{ref-mccrae1999age}{}
McCrae, R. R., Costa, P. T., Lima, M. P. de, Simões, A., Ostendorf, F.,
Angleitner, A., \ldots{} others. (1999). Age differences in personality
across the adult life span: Parallels in five cultures.
\emph{Developmental Psychology}, \emph{35}(2), 466.

@@

\hypertarget{ref-stenner2005rights}{}
Stenner, P. (2005). Rights and emotions, or: The importance of having
the right emotions. \emph{History \& Philosophy of Psychology},
\emph{7}(1), 1--11.

@@

\begin{table}[tbp]
\begin{center}
\begin{threeparttable}
\caption{\label{tab:table3}Regression Table Predicting Social Conservatism From Big-Five Personality Traits.}
\begin{tabular}{lllll}
\toprule
Predictor & \multicolumn{1}{c}{$b$} & \multicolumn{1}{c}{95\% CI} & \multicolumn{1}{c}{$t(125)$} & \multicolumn{1}{c}{$p$}\\
\midrule
Intercept & 25.82 & $[0.07$, $51.58]$ & 1.98 & .049\\
Openness to experiences & -3.73 & $[-7.50$, $0.04]$ & -1.96 & .052\\
Conscientiousness & 0.07 & $[-3.98$, $4.11]$ & 0.03 & .974\\
Extraversion & 1.56 & $[-1.42$, $4.55]$ & 1.04 & .302\\
Agreeableness & 3.93 & $[0.69$, $7.17]$ & 2.40 & .018\\
Emotional stability & 4.30 & $[0.70$, $7.91]$ & 2.37 & .020\\
\bottomrule
\addlinespace
\end{tabular}
\begin{tablenotes}[para]
\normalsize{\textit{Note.} Residual standard error: 25.13 on 125 degrees of freedom.
Multiple R-squared: 0.168, Adjusted R-squared: 0.135. F(5, 125): 5.051, p-value: 0.0003.}
\end{tablenotes}
\end{threeparttable}
\end{center}
\end{table}

\begin{table}[tbp]
\begin{center}
\begin{threeparttable}
\caption{\label{tab:table4}Regression Table Predicting Economic Conservatism From Big-Five Personality Traits.}
\begin{tabular}{lllll}
\toprule
Predictor & \multicolumn{1}{c}{$b$} & \multicolumn{1}{c}{95\% CI} & \multicolumn{1}{c}{$t(125)$} & \multicolumn{1}{c}{$p$}\\
\midrule
Intercept & 56.54 & $[34.71$, $78.37]$ & 5.13 & < .001\\
Openness to experiences & -3.79 & $[-6.98$, $-0.59]$ & -2.35 & .021\\
Conscientiousness & 1.78 & $[-1.65$, $5.20]$ & 1.03 & .307\\
Extraversion & 0.61 & $[-1.92$, $3.14]$ & 0.48 & .632\\
Agreeableness & -1.13 & $[-3.88$, $1.61]$ & -0.82 & .415\\
Emotional stability & 2.34 & $[-0.71$, $5.39]$ & 1.52 & .131\\
\bottomrule
\addlinespace
\end{tabular}
\begin{tablenotes}[para]
\normalsize{\textit{Note.} Residual standard error: 21.3 on 125 degrees of freedom. Multiple R-squared: 0.081, Adjusted R-squared: 0.044. F(5, 125): 2.207, p-value: 0.058.}
\end{tablenotes}
\end{threeparttable}
\end{center}
\end{table}
